Elements of meaningful data sharing

### 2021-03-29 new-paper.Rmd
add a note on metadata reminding using rectangular data again.
change no journals to few journals

Nosek [-@Nosek2019] describes the strategies for culture change, especially when the status quo is too entrenched. Beyond motivation, skills, tools, and incentives are critical to enable such change. In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, then increase awareness to make it normative. Once there is a critical mass, find ways to reward those who put it into practice. After these steps, it would then be possible to require this of everyone Figure 1, table 1). This framework provides a useful lens through which to examine adoption of code, computational environments and data for reproducibility. With code, we have slowly progressed through these stages. While no journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit. All of the necessary foundation to make code sharing required are in place. For computational environments, Docker makes it possible to share them in most common use cases [@Boettiger2015]. But running them still requires expertise and time on the part of reviewers. More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and slowly gaining adoption. There is still a long way to go before sharing computational environments in academia becomes mainstream. Finally, for data, we have implemented these steps out of order. The FAIR data principles were introduced in 2016 with good intentions [@Wilkinson2016] and describe the desired attributes for research data. That they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and be reusable. We as a community started out by making data required before demonstrating how this would directly benefit those who share data. Researchers are asked to deposit their data in one of many institutional or commercial repositories. Many journals even require it as a condition for publication via policy. However, the cost of preparing data for reuse remains high, and the career incentives to do so are almost non-existent. Until we make data reuse easier and more rewarding, no level of policy is going to improve the state of data sharing.

@@

Use metadata with your data. There are three benefits to a dataset having metadata: (1) Provide context by describing variables to humans; (2) Avoid errors by describing and preserving variable types for machine readability; (3) Make data easier to find online by facilitating indexing [@google-data-search]. Three pieces of metadata should be considered: (1) **variable names**, (2) **variable descriptions**, and (3) **unit text**. These should be curated by a person. Metadata created should be clearly labelled in a folder or in files named METADATA.

>>>>

Nosek [-@Nosek2019] describes the strategies for culture change, especially when the status quo is too entrenched. Beyond motivation, skills, tools, and incentives are critical to enable such change. In a five step process, Nosek describes how achieving lasting culture change requires us to show proof of concept that an activity is possible, then improve the infrastructure to make it easy, then increase awareness to make it normative. Once there is a critical mass, find ways to reward those who put it into practice. After these steps, it would then be possible to require this of everyone Figure 1, table 1). This framework provides a useful lens through which to examine adoption of code, computational environments and data for reproducibility. With code, we have slowly progressed through these stages. While few journals require that authors share code, the venues to share, along with the necessary training have paved the way for incentives and credit. All of the necessary foundation to make code sharing required are in place. For computational environments, Docker makes it possible to share them in most common use cases [@Boettiger2015]. But running them still requires expertise and time on the part of reviewers. More recent open services like Binder [@Jupyter2018] make it possible to run notebooks atop Docker images, but the technology is still new and slowly gaining adoption. There is still a long way to go before sharing computational environments in academia becomes mainstream. Finally, for data, we have implemented these steps out of order. The FAIR data principles were introduced in 2016 with good intentions [@Wilkinson2016] and describe the desired attributes for research data. That they be findable (good metadata), accessible (deposited somewhere persistent), interoperable (easily combined with other datasets), and be reusable. We as a community started out by making data required before demonstrating how this would directly benefit those who share data. Researchers are asked to deposit their data in one of many institutional or commercial repositories. Many journals even require it as a condition for publication via policy. However, the cost of preparing data for reuse remains high, and the career incentives to do so are almost non-existent. Until we make data reuse easier and more rewarding, no level of policy is going to improve the state of data sharing.

@@

Use metadata with your data. As stated at the start of the paper, we refer to tabular data. There are three benefits to a dataset having metadata: (1) Provide context by describing variables to humans; (2) Avoid errors by describing and preserving variable types for machine readability; (3) Make data easier to find online by facilitating indexing [@google-data-search]. Three pieces of metadata should be considered: (1) **variable names**, (2) **variable descriptions**, and (3) **unit text**. These should be curated by a person. Metadata created should be clearly labelled in a folder or in files named METADATA.
